Implementation and comparison of register allocation to translate \lstinline!LLVM--! to x86

### 2023-12-20 article.tex
Highlight changes

The simplest ones to implement are the \lstinline!def! and \lstinline!use! functions, as all of the values of interest are located immediately within the instruction itself and not hidden behind some layer of indirection:
\begin{minted}[linenos,bgcolor=LightGray]{ocaml}
let def (s : S.SS.t) (insn : Cfg.insn) =
  match insn with Insn (Some dop, _) -> S.SS.add dop s | _ -> s

@@

Let \(k\) be the number of working registers available on the target architecture. After building the interference graph \(G\), a node \(n\) with fewer than \(k\) neighbors is chosen. As \(n\) has at most \(k-1\) neighbors, it can be removed safely, effectively simplifying \(G\). It is pushed to a stack in order to preserve the order in which they are removed so that \(G\) can be rebuilt and the corresponding registers can be assigned correctly once a \(k\)-colorable assignment is found.
%  (i.e. \(\mathit{deg}(n) < k\))

@@

Coalescing is the process of eliminating moves/copies of data from one GPR to another by combining their interference graph nodes. This is similar to but  not the same as  the lack of interference edges between variables subject to move operations. This is because variables \(a\) and \(b\) may still be assigned different registers or even spilled if,  for instance, either of them are of significant degree. Coalescing joins nodes \(a\) and \(b\) to node \(ab\) preserving the edges of both to maintain soundness.

Since all edges are preserved, the resulting node \(an\) may be of a much higher degree. Because to this, only strategies that produce a \(k\)-colorable graph are worth considering as worth considering as additional spills negate the purpose entirely.

@@

As linear scan

@@

Validating the correctness of each instruction translation is a complicated task given how low-level it is. One approach is attaching a debugger and examining the state before and after the translated sequence of x86 instructions is executed. The result of applying a binary operation should only affect the assigned register, for instance.

@@

>>>>

\color{mygreen}
The simplest ones to implement are the \lstinline!def! and \lstinline!use! functions, as all of the values of interest are located immediately within the instruction itself and not hidden behind some layer of indirection:
\begin{minted}[linenos,bgcolor=LightGray]{ocaml}
let def (s : S.SS.t) (insn : Cfg.insn) =
  match insn with Insn (Some dop, _) -> S.SS.add dop s | _ -> s

@@

Let \(k\) be the number of working registers available on the target architecture. After building the interference graph \(G\), a node \(n\) with fewer than \(k\) neighbors is chosen. As \(n\) has at most \(k-1\) neighbors, it can be removed safely, effectively simplifying \(G\). It is pushed to a stack in order to preserve the order in which they are removed so that \(G\) can be rebuilt and the corresponding registers can be assigned correctly once a \(k\)-colorable assignment is found.

\color{black}
%  (i.e. \(\mathit{deg}(n) < k\))

@@

Coalescing is the process of eliminating moves/copies of data from one GPR to another by combining their interference graph nodes. \color{mygreen} This is similar to but  not the same as  the lack of interference edges between variables subject to move operations. This is because variables \(a\) and \(b\) may still be assigned different registers or even spilled if,  for instance, either of them are of significant degree. Coalescing joins nodes \(a\) and \(b\) to node \(ab\) preserving the edges of both to maintain soundness.

Since all edges are preserved, the resulting node \(an\) may be of a much higher degree. Because to this, only strategies that produce a \(k\)-colorable graph are worth considering as worth considering as additional spills negate the purpose entirely.
\color{black}

@@

%TODO: As linear scan

@@

\color{mygreen}
Validating the correctness of each instruction translation is a complicated task given how low-level it is. One approach is attaching a debugger and examining the state before and after the translated sequence of x86 instructions is executed. The result of applying a binary operation should only affect the assigned register, for instance.
\color{black}

@@

% TODO: uddyb mere om instrumentering og måske mere om cache meisses osv.
%
% FIXME: husk at nævn hvordan størrelsen af kompileret kode har en indflydelse på performance
